---
title: "Barcelona Counsillors' Travel Expenses"
author: "Admin"
date: 2022-05-09T21:13:14-05:00
categories: ["R"]
tags: ["Open Data BCN", "plotly", "treemap"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE)
```

## Introduction

Owing to the [Open Data BCN](https://opendata-ajuntament.barcelona.cat/es) project, Barcelona City Council provides several datasets which are almost ready for consumption with minimal pre-processing. Among them, the tables of [travel expenses of members of the town council](https://opendata-ajuntament.barcelona.cat/data/es/dataset/viatges-govern-bcn) (excluding the mayor) by year contain information about cost, duration and destination of the travels of said members since 2019. ^[Some data preparation is necessary to correct typing errors, account for the fact that some travels are split in several legs even though they have the same accounting id, etc]

The information is organized in csv files with a record for each travel accounting id (*Numero_de registre* in the files) and expense category: travel proper (*Desplaçament*), lodging (*Allotjament*) and other (*Altres despeses*). Each record contains the name of the member, city and country of destination, amount in euros and dates of departure and arrival.

## Enter the charts

With that in mind we are ready to start exploring the data. First, we plot the aggregate expenses per year (bearing in mind that for 2022 we only have data for the first quarter).

```{r datos, echo=FALSE}
load(file = "../../../resources/_gen/assets/G.Rdata")
g
```

This is quite a plain plot but, since we have data that can be related in a hierarchical (or nested) fashion (as in ); we can use the [treemap](https://plotly.com/r/treemaps/) from plotly graphing library to make a more interesting visualization. For instance, we can present the data aggregated by year and allow a drill down to an aggregation by member or the council and then show the detail of the travel, like so:

```{r fig1, echo= FALSE}
fig1
```

By clicking on the rectangles 

Another way of looking at the data is by first aggregating by council member and then by year:

```{r fig2, echo= FALSE}
fig2
```

## If everythig fails...

Sometimes it may be more convenient to just provide acces to the granular data and let the user do the queries. In this case, luckily this is quite eays thanks to the features of [reactable](https://cran.r-project.org/web/packages/reactable/index.html)

```{r rt, echo= FALSE}
rt
```